---
title: "Aggregating Hate Crime News Reports"
author: "Jeff Huang"
date: "4/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(janitor)
library(ggplot2)
library(lubridate)
library(ggthemes)

hate_crimes <- read.csv("hatecrimes2017.csv") %>% 
  clean_names() %>% 
  mutate(crime_day= as.POSIXct(article_date, format = "%m/%d/%y")) %>% 
  filter(month(crime_day) > 3)


```

# Trend Analysis
This dataset uses instances of news reports on Google News to track hate crimes from the Spring/Summer seasons in 2017. This dataset was obtained from the Propublica Data center (https://www.propublica.org/datastore/dataset/documenting-hate-news-index), compiled with Google News search results. So far, I've looked into the different trends over time, as well as looking at trends within types of hate crimes using the title and keywords columns. 


```{r data cleaning, echo=FALSE, message=FALSE}

crimes_by_day <- hate_crimes %>% 
  group_by(crime_day) %>% 
  summarize(num = n())


crimes_by_city <- hate_crimes %>% 
  select(city, state) %>% 
  drop_na(city, state) %>% 
  mutate(city_state = paste(city,state, sep=", ")) %>% 
  group_by(city_state) %>% 
  summarize(city_total = n())

lgbtq_crimes <- hate_crimes %>% 
  select(keywords, article_date, article_title) %>% 
  mutate(crime_day= as.POSIXct(article_date, format = "%m/%d/%y")) %>% 
  filter(str_detect(keywords, "gay") | str_detect(article_title, "gay")) %>% 
  group_by(crime_day) %>% 
  summarize(total = n())

iphobia_crimes <- hate_crimes %>% 
  select(keywords, article_date, article_title) %>% 
  mutate(crime_day= as.POSIXct(article_date, format = "%m/%d/%y")) %>% 
  filter(str_detect(keywords, "muslim") | str_detect(article_title, "muslim")) %>% 
  group_by(crime_day) %>% 
  summarize(total = n()) 
  

ggplot() +
  geom_line(data= crimes_by_day, aes(x=crime_day, y=num)) + 
  geom_line(data = lgbtq_crimes, aes(x=crime_day, y=total, color = "red")) + 
  geom_line(data = iphobia_crimes, aes(x=crime_day, y=total, color = "blue")) +
  labs(title = "Spring-Summer Hate Crime Activity (2017)",
       subtitle = "According to counts of Google News Reports",
       caption = "Source: Propublica",
       x = "Month",
       y = "Number of Results") +
  # formatting time labels
  scale_x_datetime(date_labels = "%B") + 
  scale_color_discrete(name = "Hate Crime Type", labels = c("Islamophia", "LGBTQ")) +


  # add theme
  theme_economist()


```

